---
title: "WestNileVirus Group"
author: "Haris Jabbar"
date: "05/09/2015"
output: html_document
---

```{r,echo=FALSE}
library(ggmap)
setwd("~/Dropbox/CSE/Summer 2015/Praktikum Data Mining/dataset/")
traindata=read.csv("train.csv")
weatherdata=read.csv("weather.csv")
spraydata=read.csv("spray.csv")
chicago_map=qmap(location="Chicago",source="google",maptype="terrain")
bbox=c(min(spraydata$Longitude),min(spraydata$Latitude),max(spraydata$Longitude),max(spraydata$Latitude))
boundingbox_map=qmap(location=bbox,source="google",maptype="terrain")
Add=unique(traindata[,c(2,8,9)])

MosquitoCountsByDate=aggregate(cbind(NumMosquitos,WnvPresent)~Date,data=traindata,sum)
MosquitoCountsByAddress_Species=aggregate(cbind(NumMosquitos,WnvPresent)~Address+Species,data=traindata,sum)
MosquitoCountsByLatLong=merge(MosquitoCountsByAddress_Species,Add)
MosquitoCountsVsWeather=merge(weatherdata,MosquitoCountsByDate)
```

# Habitation area of different species

The Habitation area of difference species is tabulated in the following table (only first 6 records are displayed)

```{r,echo=FALSE}
head(MosquitoCountsByLatLong)
```

# Cluster the Satellite Traps

The following code separates the central and satellite traps by using the length of `Trap` variable in training data.

```{r}
traps_central=traindata[nchar(as.character(traindata$Trap))!=5,]
traps_satellite=traindata[nchar(as.character(traindata$Trap))==5,]
head(traps_satellite)
```

The plot of these traps is shown in the map below

```{r}
chicago_map+geom_point(aes(y=Latitude,x=Longitude),data=traps_central, alpha=0.5,color="darkred",size=5)+
  geom_point(aes(y=Latitude,x=Longitude),data=traps_satellite, alpha=0.5,color="yellow",size=5)
```

# Plot infected counts over all mosquito species over time

The infected counts of all mosquito species over time are calculated as follows

```{r}
MosquitoCountsBySpecies_Date=aggregate(cbind(NumMosquitos,WnvPresent)~Date,data=traindata,sum)
head(MosquitoCountsBySpecies_Date)
```

The plot for Species CULEX RESTUANS is shown below

```{r,echo=FALSE}
  ggplot(data=subset(MosquitoCountsBySpecies_Date,Species=="CULEX RESTUANS"))+geom_line(aes(x=as.Date(Date),y=WnvPresent))
```

# Plot weather data and various attributes over all years into a single chart to see correlations between attributes

# Plot weather data for two stations over all years and see if they differ a lot. 

Taking Station 1 as base station, we check the deviation of Tavg of Station 2 from base station. The plot shows that there are few outliers (difference greater than 10 degrees F).

```{r}
TempDiff=as.numeric(subset(weatherdata,Station==2)[,5])-as.numeric(subset(weatherdata,Station==1)[,5])
qplot(TempDiff,binwidth=1,xlab="Temperature Difference of Station 2 from Station 1",ylab="No of Days")
```

Removing the outliers, station 2 is warmer than station 1 by 1.236 degrees on average

```{r}
qplot(TempDiff[TempDiff<10])
mean(TempDiff[TempDiff<10])
sqrt(var(TempDiff[TempDiff<10]))
```

# Calculate the likelihood of a species to carry the WNV

Aggregating the Mosquito counts by species and computing the probability, we get the following :

```{r}
MosquitoCountsBySpecies=aggregate(cbind(NumMosquitos,WnvPresent)~Species,data=traindata,sum)
MosquitoCountsBySpecies$Probability=(MosquitoCountsBySpecies$WnvPresent/MosquitoCountsBySpecies$NumMosquitos)*100
MosquitoCountsBySpecies
```

Thus CULEX PIPIENS is the most common specie carrying the WNV.

# Mosquitoes do not like wind, how does this correlate?

Plotting the Mosquito counts vs the average wind speed, we get the following plot

```{r}
qplot(as.numeric(MosquitoCountsVsWeather$AvgSpeed),MosquitoCountsVsWeather$NumMosquitos,main="Mosquito Count Vs Average Wind Speed",xlab="Average Speed",ylab="Mosquito Counts")
```

It can be seen that mosquitos do not like speeds between 50 and 100

# What influence do mild winters have?

The data is collected from Spring to Fall, thus winters is not included in the dataset.

# Information from Satellite Traps

The subset of traindata that has only satellite traps is given as :

```{r}
traps_satellite=traindata[nchar(as.character(traindata$Trap))==5,]
```



